---
title: "HW2"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
```

### Question 1

Your goal is to predict abalone age, which is calculated as the number of rings plus 1.5. Notice there currently is no `age` variable in the data set. Add `age` to the data set.

Assess and describe the distribution of `age`.

```{r,warning=FALSE,message=FALSE}
library(tidyverse)
library(tidymodels)
abalone<-read_csv('abalone.csv')
age<-abalone$rings+1.5
abalone<-abalone %>% 
  mutate(age)
abalone %>% 
  ggplot(aes(x = age)) + geom_histogram(color='red', bins = 30) +  theme_bw()
```

age is positively skewed, and a long tail to the right. Most of the age values in the data set are below 20.

### Question 2

Split the abalone data into a training set and a testing set. Use stratified sampling. You should decide on appropriate percentages for splitting the data.


```{r}
set.seed(2221)
abalone_split <- initial_split(abalone, prop = 0.80, strata = age)
abalone_train <- training(abalone_split)
abalone_test <- testing(abalone_split)
```


*Remember that you'll need to set a seed at the beginning of the document to reproduce your results.*

### Question 3

Using the **training** data, create a recipe predicting the outcome variable, `age`, with all other predictor variables. Note that you should not include `rings` to predict `age`. Explain why you shouldn't use `rings` to predict `age`.

Because the value of `age` is calculated by adding 1.5 to `rings`, there is a linear relationship between them, and the correlation coefficient is 1.

Steps for your recipe:

1.  dummy code any categorical predictors

2.  create interactions between

    -   `type` and `shucked_weight`,
    -   `longest_shell` and `diameter`,
    -   `shucked_weight` and `shell_weight`

3.  center all predictors, and

4.  scale all predictors.


```{r,warning=FALSE,message=FALSE}
abalone_train<-abalone_train[,-9]
abalone_recipe <- recipe(age ~ ., data = abalone_train) %>% 
  step_dummy(all_nominal_predictors())

abalone_recipe_model<-abalone_recipe %>%
  step_interact(terms = ~ shucked_weight:starts_with("type")) %>%
  step_interact(terms = ~ longest_shell:diameter) %>%
  step_interact(terms = ~ shucked_weight:shell_weight)  %>% 
  step_center(all_numeric(), -all_outcomes(), -has_role('id variable')) %>% 
  step_scale(all_numeric(), -all_outcomes(), -has_role('id variable')) 
abalone_train_1<-abalone_recipe_model %>% 
  prep() %>% 
  bake(new_data = abalone_train)

```


You'll need to investigate the `tidymodels` documentation to find the appropriate step functions to use.

### Question 4

Create and store a linear regression object using the `"lm"` engine.

```{r}
lm_model <- linear_reg() %>% 
  set_engine("lm")
```

### Question 5

Now:

1.  set up an empty workflow,
2.  add the model you created in Question 4, and
3.  add the recipe that you created in Question 3.

```{r,warning=FALSE,message=FALSE}
lm_wflow <- workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(abalone_recipe_model)
```

### Question 6

Use your `fit()` object to predict the age of a hypothetical female abalone with longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1.

```{r}
data<-data.frame(type="F",longest_shell=0.5,diameter = 0.10,
                 height = 0.30, whole_weight = 4,
                 shucked_weight = 1, viscera_weight = 2,
                 shell_weight = 1)
data<-tibble(data)
lm_fit <- fit(lm_wflow, abalone_train)
predict(lm_fit, new_data = data)
```

### Question 7

Now you want to assess your model's performance. To do this, use the `yardstick` package:

1.  Create a metric set that includes *R^2^*, RMSE (root mean squared error), and MAE (mean absolute error).
2.  Use `predict()` and `bind_cols()` to create a tibble of your model's predicted values from the **training data** along with the actual observed ages (these are needed to assess your model's performance).
3.  Finally, apply your metric set to the tibble, report the results, and interpret the *R^2^* value.

```{r}

diamond_metrics <- metric_set(rmse, rsq, mae)

abalone_train_res <- predict(lm_fit, new_data = abalone_train %>% select(-age))
abalone_train_res <- bind_cols(abalone_train_res, abalone_train %>% select(age))

diamond_metrics(abalone_train_res, truth = age, 
                estimate = .pred)
```


In the multiple linear regression, 57% of age can be determined by explain these variables.

